Hierarchical storage management in user space for neuroimaging applications

### 2025-10-14 paper/sea-neuro/paper-neuro.tex
Update paper and generated figures

\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

% \usepackage{todonotes}
% \usepackage{float}
% \newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}

% \linenumbers

@@

@@

@@

    Sea (Figure~\ref{fig:seaneuro:diagram}) is a data-management library that leverages
    the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
    glibc, on Linux systems. This enables Sea to redirect write calls
    aimed at slower storage devices to a faster device whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the details provided in an
    initialization file called \texttt{sea.ini}. This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority. Importantly, Sea is not a filesystem, but rather a
    lightweight tool that redirects files to more optimal storage devices.

    We define all locations in which Sea can redirect writes to as a cache. Such
    caches include any available storage located on the compute nodes, such as
    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs).

    As neuroimaging pipeline results typically require post-processing and
    HPC compute-local resources are only accessible during the reserved
    duration, Sea provides functionality to flush and evict data to persistent
    shared storage. To avoid interrupting ongoing
    processing with data management operations, this is accomplished via a separate thread (known as the
    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
    that need to be persisted to storage within a file called
    \texttt{.sea\_flushlist}, and temporary files which can be removed from
    cache within a file called the \texttt{.sea\_evictlist}. Both these files
    can be populated using regular expressions to denote the paths to flush and
    evict. If a file occurs in both the \texttt{.sea\_flushlist} and
    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
    operation and simply move the file from local to persistent storage rather
    than copying it. Files that will be reused by the application should only be
    flushed rather than flushed and evicted, as files can benefit from speedups
    related to reading from fast rather than slow storage. Sea currently also
    provides a rudimentary prefetch thread that can move files located within
    Sea to the fastest available cache. To use Sea's prefetch
    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
    populated using regular expressions like the flushing and eviction files.

    To interact with Sea, a mountpoint is used. The mountpoint
    is an empty directory that behaves as a view to all the files and
    directories stored within Sea. In order to keep track of the locations of
    files within the mountpoint, Sea mirrors the directory structure of each
    storage location across all caches. It is generally advisable to provide
    empty storage locations for Sea to use as the mirroring of large directories
    can take some time.

@@

\begin{figure*}[ht]
\centering

\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/slashbin_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/slashbin_withzoom.pdf}
    \caption{Zoomed between 0 and 5000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on controlled dedicated cluster. Makespan
denotes the total time between execution launch and completion of the last computing task.}
\label{fig:seaneuro:slashbin}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.8\textwidth]{figures/beluga_sea_tmpfs.pdf}%
\caption{Makespan comparison between Sea and tmpfs on the production cluster with flushing disabled}
\label{fig:seaneuro:tmpfs}
\end{figure*}

@@

\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafull}
\end{subfigure}
\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled}
\label{fig:seaneuro:beluga-noflush}
\end{figure*}

@@

\begin{figure*}[ht]
  \centering
\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{0.8\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/beluga_withflush_zoom.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing enabled for all files}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}

>>>>

\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
% \usepackage{todonotes}
% \usepackage{float}
% \newcommand{\TG}[1]{\todo[color=green!60!black]{\textbf{\textsc{From Tristan:}} #1\xspace}}
\newcommand{\vhs}[1]{\textcolor{red}{#1}}
\linenumbers

@@

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{myyamlstyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

@@

    Sea (\autoref{fig:seaneuro:diagram}) is a data-management library that leverages
    the \texttt{LD\_PRELOAD}
    trick to intercept POSIX file system calls, or more specifically, file access calls to the GNU C library,
    glibc, on Linux systems. Its goal is to easily enable users to intergrate efficient data management within their
    applications within their existing workflows. Using glibc interception, Sea can redirect write calls
    aimed at slower storage devices to a faster device whenever possible.
    Similarly, when intercepting read calls, Sea can choose to read from a
    faster device if a copy is available on that device. Sea decides which
    storage location it can write to based on the details provided in an
    initialization file called \texttt{sea.ini} (\autoref{lst:seaini}). This file informs Sea of which
    locations it can use to read and write to, and well as their order of
    priority. Importantly, Sea is not a filesystem, but rather a
    lightweight tool that redirects files to more optimal storage devices.

    \begin{figure}
      \centering
       \begin{lstlisting}[style=myyamlstyle, caption={Example Sea configuration file (sea.ini)}, label=lst:seaini]
[Sea]                                                                  
mount_dir = /lustre/seamount ;                                    
n_levels = 3 ;                                                        
cache_0 = /dev/shm/seasource ;                                           
cache_1 = /localscratch/seasource ;                                         
cache_2 = /lustre/seasource ;                                    
log_level = 0 ;                                 
log_file = /lustre/sea.log ;                             
max_fs = 646971392 ;                                                   
n_threads = 16 ;    \end{lstlisting}
    \end{figure}

    \vhs{To configure Sea, a few parameters need to be configured within a \texttt{sea.ini} file located in the directory from where the application will be launched. To interact with Sea, a mountpoint is used. The mountpoint
    is an empty directory that behaves as a view to all the files and
    directories stored within Sea. In order to keep track of the locations of
    files within the mountpoint, Sea mirrors the directory structure of each
    storage location across all caches. It is generally advisable to provide
    empty storage locations for Sea to use as the mirroring of large directories
    can take some time.}

    \vhs{We define all locations in which Sea can redirect reads and  writes to as a cache. Such
    caches include any available storage located on the compute nodes, such as
    tmpfs (i.e. a temporary file system in virtual memory) or solid state drives (SSDs). When the application attempts to write
    data to the Sea mountpoint, Sea will write the data to one of these cache locations, by order of preference (cache\_0 being the most optimal).
    Since many cache locations may be available, the user must specify the number of cache levels available (n\_levels) followed by the locations for each of these levels.}

    \vhs{Optionally, logging may be enabled, giving insight the glibc commands that were executed and how the filepaths were treated by Sea. There are five log levels
    available (0:None, 1:Error, 2: Warning, 3: Info, 4: Debug).\autoref{fig:seaneuro:logs} shows how logs can be interpreted to gain insight into the decisions made by Sea.}

    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of partitioned files across storage devices. Rather, it relies on configuration parameters
    \texttt{max\_fs} and \texttt{n\_threads} to determine whether there is sufficient space on a storage device to write the file to. \texttt{max\_fs} represents an upper bound estimate of the maximum
    filesize generated by the application in bytes. }

    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{figures/ex_log.png}
       \captionsetup{justification=centering}
    \captionof{figure}{Example logs outputted at log level 3 (INFO)}
    \label{fig:seaneuro:logs}
    \end{figure}

    As neuroimaging pipeline results typically require post-processing and
    HPC compute-local resources are only accessible during the reserved
    duration, Sea provides functionality to flush and evict data to persistent
    shared storage. To avoid interrupting ongoing
    processing with data management operations, this is accomplished via a separate thread (known as the
    ``flusher'') that moves data from the caches to long-term storage. Users must inform Sea of files
    that need to be persisted to storage within a file called
    \texttt{.sea\_flushlist}, and temporary files which can be removed from
    cache within a file called the \texttt{.sea\_evictlist}. Both these files
    can be populated using regular expressions to denote the paths to flush and
    evict (e.g., ~\autoref{fig:seaneuro:flushlist}). If a file occurs in both the \texttt{.sea\_flushlist} and
    \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move}
    operation and simply move the file from local to persistent storage rather
    than copying it. Files that will be reused by the application should only be
    flushed rather than flushed and evicted, as files can benefit from speedups
    related to reading from fast rather than slow storage. Sea currently also
    provides a rudimentary prefetch thread that can move files located within
    Sea to the fastest available cache. To use Sea's prefetch
    capabilities, a file called \texttt{.sea\_prefetchlist} needs to be
    populated using regular expressions like the flushing and eviction files.
    When used, the prefetch, flush and eviction configuration files must all be located within the directory from where the application will be launched.

    \begin{figure}
      \centering
      \includegraphics[width=0.5\textwidth]{figures/ex_flushlist.png}
       \captionsetup{justification=centering}
    \captionof{figure}{Example \texttt{.sea\_flishlist} used for AFNI processing}
    \label{fig:seaneuro:flushlist}
    \end{figure}

@@

@@

\begin{figure*}
\centering

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/slashbin_nozoom_updated.pdf}%
    \caption{Complete results}\label{fig:seaneuro:slashbinfull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/slashbin_withzoom_updated.pdf}
    \caption{Zoomed between 0 and 3000 seconds}\label{fig:seaneuro:slashbinzoom}
\end{subfigure}
\caption{Makespan comparison between Sea and baseline (I/O to and from Lustre) on controlled dedicated cluster. Makespan
denotes the total time between execution launch and completion of the last computing task. Five repetitions were recorded for each experimental condition (including with and without busy writers)
using Sea and baseline. The black dots (without busy writers) represent pipeline makespan when pipeline has exclusive access to the cluster.
a) shows that using Sea on SPM processing of large 
datasets produces important speedups (marker below the y=x line) when compared to baseline, due to its data intensive nature.
However, as seen in (b), for short application makespans, Sea performance is comparable to baseline. }
\label{fig:seaneuro:slashbin}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{figures/beluga_sea_tmpfs_updated.pdf}%
\caption{Makespan comparison between Sea and tmpfs (baseline) on the production cluster with flushing disabled. Mean pipeline execution time with Sea was recorded across five repetitions and compared
to time it took to execute the same pipeline exclusively using data stored on tmpfs.
As shown by the figure, Sea's mean runtime is comparable to that of writing directly to RAM, across pipelines, datasets and parallelism.}
\label{fig:seaneuro:tmpfs}
\end{figure*}

@@

\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom_updated.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafull}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\textwidth]{figures/beluga_withoutflush_nozoom_updated.pdf}
    \caption{Zoomed between 0 and 3000 seconds}\label{fig:seaneuro:belugazoom}
\end{subfigure}
\caption{Makespan comparison between Sea and Baseline on the production cluster with flushing disabled. }
\label{fig:seaneuro:beluga-noflush}
\end{figure*}

@@

\begin{figure*}[t]
  \centering
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\linewidth}
    \includegraphics[width=\columnwidth]{figures/beluga_withflush_nozoom_updated.pdf}%
    \caption{Complete results}\label{fig:seaneuro:belugafullwf}
\end{subfigure}
\begin{subfigure}{\textwidth}
    \centering
    \captionsetup{width=\textwidth}
    \includegraphics[width=\textwidth]{figures/beluga_withflush_zoom_updated.pdf}
    \caption{Zoomed between 0 and 2000 seconds}\label{fig:seaneuro:belugazoomwf}
\end{subfigure}
\caption{Makespan comparison between Sea and baseline (Lustre) on the production cluster with flushing enabled for all files.}
\label{fig:seaneuro:beluga-wflush}
\end{figure*}
